---
title: "Practical Machine Learning Writeup"
author: "Aniruddha Das"
date: "Sunday, August 24, 2014"
output: html_document
---

Read in the data to identify features that could help with classifying the excercise **classe**.
```{r}
getwd()
setwd("C:/Users/and/Documents/GitHub/PracticalMachineLearning")
dir()
train <- read.csv2("pml-training.csv",sep = ",")
head(train)
```

seems there are lots of NAs, empty values and #DIV/0! (excel division errors in the data).Lets replace all the bad values with NAs. 
```{r}
train <- read.csv2("pml-training.csv",sep = ",", na.strings = c("", "NA", "#DIV/0!") ) 
head(train)
```

Based on looking at the data seems there are columns that are not useful for prediction like X, user_name, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp, new_window and num_window.

Subsetting for relevant features names, removing NA columns and casting all entries as numeric except classe:
```{r}
validTrainFeatures <- colnames(train[colSums(is.na(train)) == 0])[-(1:7)]
validTrainData <- train[validTrainFeatures]
head(validTrainData)
complete <- function(x) {x[,sapply(x, function(y) !any(is.na(y)))] }
validTrainData <- complete(validTrainData)
for(i in c(2:ncol(validTrainData)-1)) {validTrainData[,i] = as.numeric(as.character(validTrainData[,i]))}
str(validTrainData)
```
Creating train and test sets from the valid training data. 70% training and 30% testing.
```{r}
library(caret)
inTrain <- createDataPartition(y=validTrainData$classe, p=0.7, list=FALSE)
validTrainDataTrain <- validTrainData[inTrain,]
validTrainDataTest <- validTrainData[-inTrain,]
```

Creating the random forest model and testing it on train and test data:
```{r}
library(randomForest)
random.forest <- train(validTrainDataTrain[,-53],
                       validTrainDataTrain$classe,
                       tuneGrid=data.frame(mtry=3),
                       trControl=trainControl(method="none")
                    )
predictionsTrain <- predict(random.forest, newdata=validTrainDataTrain)
confusionMatrix(predictionsTrain,validTrainDataTrain$classe)

predictionsTest <- predict(random.forest, newdata=validTrainDataTest)
confusionMatrix(predictionsTest,validTrainDataTest$classe)
```
We get very good Accuracy on the test data as shown above in the test data.